Possible kinematics of an atypical Lyman-$\alpha$ dwarf galaxy

### 2017-02-28 paper/letter.tex
intro
title intro

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[An atypical Lyman-$\alpha$ dwarf galaxy]{
Clumpiness and rotation: possible causes for an atypical Lyman-$\alpha$ dwarf galaxy}
\author[J.E. Forero-Romero et al.]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Max Gronke$^2$, 
Mar\'ia Camila Remolina-Guti\'errez$^1$,
\newauthor
Nicol\'as Garavito-Camargo$^3$, 
Mark Dijkstra$^2$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
$^2$ Institute of Theoretical Astrophysics, University of Oslo,
Postboks 1029 Blindern, NO-0315 Oslo, Norway.\\
$^3$ Department of Astronomy, University of Arizona, 933 North Cherry
Avenue, Tucson, AZ 85721, USA. 
}

@@

\begin{abstract}
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ line emission that had evaded theoretical
interpretation so far. 
In this paper we explain these features by two different models: 
an interstellar medium composed by outflowing clumps with additional
random motions and an homogeneous gaseous sphere undergoing bulk rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these physical conditions.
We find that both models independently require high velocities
(either a  a clump velocity dispersion of \sigmaclump\ with outflows of
\inftyclump\ or a bulk rotation of $348^{+75}_{-48}$ km s$^{-1}$)
consistent with a dynamical mass of at 
least a $10^{10}$ M$_{\odot}$, ten times larger than its baryonic mass.   
We argue that a possible explanation for this excess of
dynamical mass is the presence of a supermassive black hole at the
center of \tol. 
This work demonstrates the importance of considering multiphase
physics and rotation among the possible conditions shaping the
\lya\ spectra of the first galaxies.  
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of supermassive black
holes.  
\end{abstract}

@@

In this paper we show how the \tol's \lya\ profile can be explained
either by  or the recently developed class of complex multiphase models 
\citep{Gronke2016} or by a simple rotation model
\citep{GaravitoCamargo2014}. 
\section{Observations}

\tol's basic observational characteristics are summarized in Table \ref{obstable}.
Its receding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (with the Hubble constant $H_{0}$=73
Mpc km s$^{-1}$)

@@

We compare each resulting spectra to the observational results from
\tol\ after normalizing the observed and simulated spectra to have a a
flux integral of one.
We then build a $\chi^2$ on the normalized flux measurements for each one
of the $2500$ models, and then select for further analysis the best
$1\%$ models according to the lowest $\chi^2$ values.
The
$\chi^2$ gap in those $25$ models is close to $3000$, the lowest
$\chi^2$ being close to $1200$.
The total number of degrees of freedom is $104$.

Then we performed a Kolmogorov-Smirnov test to compare each parameter
distribution in the best $25$ models against the parent distribution
of $2500$ models. 
If we obtain a p-value $<0.05$ for a given model parameter, we conclude that
this parameter does influence the $\chi^2$ fit, as the distribution for
the best $\chi^2$ models is statistically different to the
distribution from the global sample of $2500$ models.

The best values for the influential parameters correspond to
the values that produce the minimum
$\chi^2$.
The 1-$\sigma$ uncertainty comes from a parabolic fit to the
$\chi^2$ as a function of $v_{\infty,\rm{cl}}$, $\sigma_{\rm cl}$,
$P_{\rm cl}$ around its corresponding minimum.

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[An atypical Lyman-$\alpha$ dwarf galaxy]{
Possible kinematics of an atypical Lyman-$\alpha$ dwarf galaxy}
\author[J.E. Forero-Romero et al.]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Max Gronke$^2$, 
Mar\'ia Camila Remolina-Guti\'errez$^1$,
\newauthor
Nicol\'as Garavito-Camargo$^3$, 
Mark Dijkstra$^2$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
$^2$ Institute of Theoretical Astrophysics, University of Oslo,
Postboks 1029 Blindern, NO-0315 Oslo, Norway.\\
$^3$ Department of Astronomy, University of Arizona, 933 North Cherry
Avenue, Tucson, AZ 85721, USA. 
}

@@

\begin{abstract}
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ line emission that had evaded theoretical
interpretation so far. 
In this paper we explain these features by two different models: 
an interstellar medium composed by outflowing clumps with additional
random motions and an homogeneous gaseous sphere undergoing bulk rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
We find that both models independently require high velocities
(either a  a clump velocity dispersion of \sigmaclump\ with outflows of
\inftyclump\ or a bulk rotation of $348^{+75}_{-48}$ km s$^{-1}$)
consistent with a dynamical mass of at 
least a $10^{10}$ M$_{\odot}$, ten times larger than its baryonic mass.   
We argue that a possible explanation for this excess of
dynamical mass is the presence of a supermassive black hole at the
center of \tol. 
This work demonstrates the importance of considering multiphase
kinematics and rotation among the possible conditions shaping the
\lya\ spectra of the first galaxies.  
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of supermassive black
holes.  
\end{abstract}

@@

In this paper we show how the \tol's \lya\ profile can be explained
either by  or the recently developed class of complex multiphase models 
\citep{Gronke2016} or by a simple rotation model
\citep{GaravitoCamargo2014}. 
The models we use correspond to simplified geometrical configurations
to allow a deep exploration of parameter space and gain, for the firs time,
some physical insight into the \tol's kinematic properties.

\section{Observations}
\tol's basic observational characteristics are summarized in Table \ref{obstable}.
Its receding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (with the Hubble constant $H_{0}$=73
Mpc km s$^{-1}$)

@@

We compare each resulting spectra to the observational results from
\tol\ after normalizing the observed and simulated spectra to have a a
flux integral of one.
We then build a $\chi^2$ on the normalized flux measurements for each one
of the $2500$ models, and then select for further analysis the best
$1\%$ models according to the lowest $\chi^2$ values.
The
$\chi^2$ gap in those $25$ models is close to $\Delta\chi^2 = 3000$, the lowest
$\chi^2$ being close to $\chi^2_{min}=1200$.
The total number of degrees of freedom is $104$.

Then we perform a Kolmogorov-Smirnov test to compare each parameter
distribution in the best $25$ models against the parent distribution
of $2500$ models. 
If we obtain a p-value $<0.05$ for a given model parameter, we conclude that
this parameter does influence the $\chi^2$ fit, as the distribution for
the best $\chi^2$ models is statistically different to the
distribution from the global sample of $2500$ models.

The best values for the influential parameters correspond to
the values that produce the minimum
$\chi^2$.
The 1-$\sigma$ uncertainty comes from a parabolic fit to the
$\chi^2$ as a function of the influential parameters around its
corresponding minimum.  
%$v_{\infty,\rm{cl}}$, $\sigma_{\rm cl}$,
%$P_{\rm cl}$
